Anony\emph{mice}d shareable data: Using \emph{mice} to create and analyze multiply imputed synthetic data sets

### 2021-10-27 manuscript/Manuscript.tex
Incorporate reviewer comments.

\hypertarget{add-reviewer-name}{%
\section{ADD REVIEWER NAME}\label{add-reviewer-name}}

@@

An alternative solution has been proposed separately by
\citet{rubin_statistical_disclosure_1993} and
\citet{little_statistical_1993}. Although their approaches differ to
some extent, the overarching procedure is to use bonafide observed data
to generate multiply imputed synthetic data sets that can be freely
disclosed. While in practice, one could see this as replacing the
observed data values by multiple draws from the posterior predictive
distribution of the observed data, based on some imputation model, Rubin
would argue that these synthetic data values are merely draws from the
same true data generating model. In that sense, the observed data is
never replaced, but the population is resampled from the information
captured in the (incomplete) sample. Using this approach, the researcher
could replace the observed data set as a whole with multiple synthetic
versions. Alternatively, the researcher could opt to only replace a
subset of the observed data. For example, one can choose to only replace
dimensions in the data that could be compared with publicly available
data sets or registers. Likewise, synthetisation could be limited to
those values that are disclosive, such as high incomes or high
turnovers.

Conceptually, the synthetic data framework is based upon the building
blocks of multiple imputation of missing data, as proposed by
\citet{rubin_multiple_1987}. Instead of replacing just the missing
values with multiple draws from the posterior predictive distribution,
one could easily replace any observed sensitive values. Similarly to
multiple imputation of missing data, the multiple synthetic data sets
allow for correct statistical inferences, despite the fact that the
analyses do not use the ``true'' value. The analyses over multiple
synthetic data sets should be pooled into a single inference, so that
the researcher can draw valid conclusions from the pooled results. To
that respect, the variance should reflect the added variability that is
induced by the imputation procedure.

Potentially, this approach could fulfill the needs for openly accessibly
data, without running into barriers with regard to privacy and
confidentiality constraints. However, there is no such thing as a free
lunch: data collectors have to put effort in creating high-quality
synthetic data. The quality of the synthetic data is highly dependent on
the imputation models, and using flawed models to generate synthetic
data might bias subsequent analyses
\citep{reiter2004simultaneous, grund2021using, jiang2021balancing}.
Conversely, if the models used to create the synthetic data are able to
preserve the relationships between the variables as in the original
data, the synthetic data can be nearly as informative as the observed
data. Thus, to fully exploit the benefits of synthetic data, additional
complications to actually create these high-quality data sets should be
kept at a minimum.

@@

The \texttt{mice} package \citep{mice} in \texttt{R} \citep{Rproject}
has been developed for multiple imputation to overcome problems related
to nonresponse. In that context, the aim is to replace missing values
with plausible values from the posterior predictive distribution of that
variable. In \texttt{mice}, this is accomplished using fully conditional
specification (FCS) \citep{vanbuuren_fully_2006}, which breaks down the
multivariate distribution of the data
\(\textbf{Y} = (\textbf{Y}_{obs}, \textbf{Y}_{mis})\) into
\(j = 1, 2, \dots, k\) univariate conditional densities, where \(k\)
denotes the number of columns in the data. Using FCS, a model is
constructed for every incomplete variable and the missing values
\(Y_{j, mis}\) are then imputed with draws from the posterior predictive
distribution of \(P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)\) on a
variable-by-variable basis. Note that the predictor matrix \(Y_{-j}\)
may contain yet imputed values from an earlier imputation step, and thus
will be updated after every iteration. This procedure is applied \(m\)
times, resulting in \(m\) completed data sets
\(\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})\),
with \(\textbf{D}^{(l)} = (\textbf{Y}_{obs}, Y^{(l)}_{mis})\).

In \texttt{mice}, the generation of multiply imputed data sets to solve
for unobserved values is straightforward. The following pseudocode
details the multiple imputation of the \texttt{mice::boys} data set
\citep{fredriks_boys_2000} into the object \texttt{imp} with
\texttt{m\ =\ 10} imputated sets and \texttt{maxit\ =\ 7} iterations for
the algorithm to converge, using the default imputations methods given
each column's data class.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(mice)}
\NormalTok{imp }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(boys, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{7}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

It is straightforward to extend the imputation approach to generate
synthetic values. Rather than imputing missing data, the observed values
are then replaced by synthetic draws from the posterior predictive
distribution. For simplicity, assume that the data is completely
observed (i.e., \(\textbf{Y} = \textbf{Y}_{obs}\)). Following the
notation of \citet{reiter_raghunathan_multiple_2007}, let, given \(n\)
observations, \(Z_i = 1\) if any of the values of unit
\(i = 1, 2, \dots, n\), are to be replaced by imputations, and
\(Z_i = 0\) otherwise, with \(\textbf{Z} = (Z_1, Z_2, \dots, Z_n)\).
Accordingly, the data consists of values that are to be replaced and
values that are to be kept (i.e.,
\(\textbf{Y} = (\textbf{Y}_{rep}, \textbf{Y}_{nrep})\). Now, instead of
imputing \(\textbf{Y}_{mis}\) with draws from the posterior predictive
distribution of \(P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)\) as in the
missing data case, \(\textbf{Y}_{rep}\) is imputed from the posterior
distribution of
\(P(Y^{(l)}_{j, rep} | \textbf{Y}^{(l)}_{-j}, \textbf{Z}, \theta)\),
where \(l\) is an indicator for the synthetic data set
(\(l = 1, 2, \dots, m\)). The synthetic values that are imputed at an
earlier step can be used for imputing variable \(j\). This process
results in the synthetic data
\(\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})\).
Note that, given a completely observed data set, iterating over the
posterior predictive distribution is not required.

If the goal is to create synthetic versions of a data set that suffers
from missingness, the imputation procedure becomes more complex. A
two-step imputation procedure has been proposed by
\citet{reiter2004simultaneous}. In the first step, the missing data is
imputed according to the principles of multiple imputation for missing
data, while in a second step, the synthetic data is imputed conform the
corresponding procedure. Intuitively, these steps could be combined
using \texttt{mice}, conditioning the imputations on both the
missingness and the values that are to be synthesized. In practice,
replacing observed and missing cells in the \texttt{mice::boys} data set
by synthetic values into the object \texttt{syn}, given the same
imputation parameters as before, can be realized by the following code
execution.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{syn }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(boys, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{7}\NormalTok{, }
            \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{, }
                           \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(boys),}
                           \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(boys)))}
\end{Highlighting}
\end{Shaded}

Here, the argument \texttt{where} requires a matrix of the same
dimensions as the data, (i.e., a \(n \times k\) matrix) containing
logicals \(z_{ij}\) that indicate which cells are selected to have their
values replaced by draws from the posterior predictive distribution.
This approach allows to replace a subset of the observed data, or as in
the aforementioned example, the observed data as a whole, resulting in a
data set that partially or completely consists of synthetic data values.
However, because the validity of this approach when the data suffers
from missingness has not been investigated yet, we restrict ourselves in
the following sections to a completely observed data set.However,
because the validity of this approach has not been investigated yet, we
restrict ourselves in the followi

Choosing an adequate imputation model to impute the data is paramount,
as a flawed imputation model may drastically impact the validity of
inferences \citep{grund2021using, jiang2021balancing}. Imputation models
should be as flexible as possible to capture most of the patterns in the
data, and to model possibly unanticipated data characteristics
\citep{murray_multiple_2018, rubin_18years_1996}. Parametric methods,
albeit easy to implement in practice, may be too restrictive to capture
generally complex patterns in the data, especially in the case of
nonlinear relations and interactions between multiple variables.
Classification and regression trees \citep[CART;][]{breiman_cart_1984}
allow to model more complex patterns in the data, and have therefore
been suggested as an appropriate imputation method
\citep{reiter_cart_2005, burgette_reiter_cart_2010, doove_buuren_recursive_2014}.
Loosely speaking, CART sequentially splits the predictor space into
non-overlapping regions in such a way that the within-region variance is
as small as possible after every split. As such, CART does not impose
any parametric distribution on the data, making it a widely applicable
method that allows for a large variety of relationships within the data
\citep{islr_2013}. Given these appealing characteristics and the call
for the use of flexible methods when multiply imputing data, we will
focus our illustrations and evaluations of \texttt{mice} to method
\texttt{mice.impute.cart()}, realized by:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{syn }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(boys, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{7}\NormalTok{, }
            \AttributeTok{method =} \StringTok{"cart"}\NormalTok{,}
            \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{, }
                           \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(boys),}
                           \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(boys)))}
\end{Highlighting}
\end{Shaded}

In a nutshell, the above code shows the simplicity of creating
\(m = 10\) synthetic data sets using \texttt{mice}. In practice,
however, one should take some additional complicating factors into
account. For example, one should account for deterministic relations in
the data. Additionally, relations between variables may be described
best using a different model than \texttt{CART}. Such factors are data
dependent and should be considered by the imputer. In the next section,
we will describe how the \texttt{boys} data can be adequately imputed.
Additionally, we will show through simulations that this approach yields
valid inferences.

@@

Usually, one would draw samples from a population that can be
synthesized to evaluate the performance of the synthesization methods.
As we only have access to a single sample, 1000 bootstrap samples of the
\texttt{boys} data have been synthesized with \(m = 5\) imputations for
every data cell to induce an appropriate amount of sampling variance.
Synthetic values are generated using the \texttt{CART} imputation method
for all columns, except for \texttt{bmi}. The deterministic relation
\texttt{bmi} which will be synthesized passively based on the synthetic
values for \texttt{hgt} and \texttt{wgt} to preserve the relation in the
synthetic data. Additional parameters that come with the use of
\texttt{mice.impute.cart()} are the complexity parameter \texttt{cp} and
the minimum number of observations in any terminal node
\texttt{minbucket}, that both constrain the flexibility of the
imputation model. The values of the parameters \texttt{cp} and
\texttt{minbucket} ought to adhere to the call for imputation models
that are as flexible as possible. Appropriate values for these
parameters, as well as the input for the \texttt{predictorMatrix},
depend on the data at hand. In the current example, the complexity
parameter is specified at \texttt{cp\ =\ 1e-08} rather than the default
value \texttt{1e-04}, and the minimum number of observations in each
terminal node is set at \texttt{minbucket} \(= 3\) rather than the
default value \(5\). By allowing for more complexity in the imputation
model, bias in the estimates from the synthetic data set is reduced.
Additionally, since the synthesis pattern is monotone, the number of
iterations can be set to \texttt{maxit\ =\ 1} \citep[e.g.,][Ch.
3]{drechsler_synthetic_2011}.

@@

Obtaining a final point estimate of the parameter of interest \(Q\)
after imputation is fairly easy and no different from pooling in the
case of missing data \citep{rubin_multiple_1987}. One can calculate the
average of the \(m\) point estimates \(q^{(l)}\) \[
\bar{q}_m = \sum_{l = 1}^m \frac{q^{(l)}}{m},
\] with \(l = 1, \dots, m\).

Also similarly to the missing data case, variances, and subsequently
confidence intervals, should incorporate the increase in variance that
is due to imputation
\citep{reiter_partially_inference_2003, drechsler_synthetic_2011}. Yet,
the increase in variance due to imputation differs according to whether
missing values are imputed or observed data is replaced by synthetic
values. Whereas the variance estimate after imputation of missing data
needs to account for the fact that a certain amount of information in
the data is missing, variance estimation from synthetic data does not
suffer from this issue. The adjusted variance estimate that follows from
using multiple synthetic data sets only suffers from the fact that a
finite number of \(m\) synthetic data sets are used to resemble the
observed data. Hence, the according variance estimate for synthetic data
as developed by \citet{reiter_partially_inference_2003} yields \[
T = \bar{u}_m + \frac{b_m}{m},
\]

@@

\[
\bar{u}_m = \sum_{l = 1} \frac{u^{(l)}}{m},
\] where \(u^{(l)}\) denotes the variance estimate in the \(l\)th
synthetic data set.

@@

The univariate descriptives for the original data and the synthetic data
can be found in Table 2.

@@

We demonstrate that generating synthetic data sets with \texttt{mice} in
\texttt{R} is a straightforward process that fits well in a data
analysis pipeline. The approach is hardly different from using multiple
imputation to solve problems related to missing data, and hence can be
expected to be familiar to applied researchers. The multiple synthetic
sets yield valid inferences on the true underlying data generating
mechanism, thereby capturing the nature of the original data. This makes
the multiple synthetisation procedure with \texttt{mice} suitable for
further dissemination of synthetic data sets.

To some, the procedure of generating multiple synthetic sets may seem
overly complicated. We would like to emphasize that analyzing a single
synthesized set, while perhaps unbiased, would underestimate the
variance properties that are so important in drawing statistical
inferences from finite data sets. After all, we are often not interested
in the sample at hand, but aim to make inferences about the underlying
data generating mechanism as reflected in the population. Properly
capturing the uncertainty of synthetic data sets, just like with
incomplete data sets, is therefore paramount.

Besides capturing uncertainty of synthetic data, it is important that
imputers pay close attention to disclosure risks that remain after
creating synthetic data sets. Creating synthetic data, unless generated
from a completely parametric distribution, does not remove all potential
disclosure risks. For example, if the values that ought to be replaced
get the exact same value imputed, the synthetisation procedure has no
use.

Additionally, if not all variables in the data are synthesized, but the
variables that are synthesized can be linked to open access data, a
disclosure risk may remain \citep{drechsler2011empirical}. If the open
access data allows for identification, and the corresponding
observations in the synthetic data can be identified, the variables that
are not synthesized may provide information that should not have been
disseminated. The associated problems generally decrease when the
sensitive variables are synthesized as well. Still, it is important to
remain critical regarding the extent to which the synthetic data might
be disclosive. The practical development of easy to use software to
identify which observations are at risk of disclosure is an area of
future work that can improve these issues. Subsequently, the
implementation of ways to overcome such problems once detected, for
example by record swapping as suggested by \citet{jiang2021balancing},
is welcomed.

That said, while the fields of data disclosure control and data
synthetisation originated decades ago, it is more relevant today than
ever. In our study we outline a standard research workflow where the
goal is to synthesize a complete data set. To simplify the simulation we
adopted a bootstrapping scheme to induce sampling variance in order to
conform to the combination procedure defined by
\citet{reiter_partially_inference_2003}. Ideally, one would like to omit
the bootstrap from the synthetisation to adopt a scheme where the
sampled data itself serves as a reference: much like the procedure
outlined in \citet{vink2014pooling} for incomplete data simulation. The
corresponding pooling rules have not been derived yet and the
incorporation in data analysis workflows would require proper attention
from developers alike.

>>>>

@@

An alternative solution has been proposed separately by
\citet{rubin_statistical_disclosure_1993} and
\citet{little_statistical_1993}. Although their approaches differ to
some extent, the overarching procedure is to use bonafide observed data
to generate multiply imputed synthetic data sets that can be freely
disclosed. While in practice, one could see this as replacing the
observed data values by multiple draws from the predictive distribution
of the observed data, based on some imputation model, Rubin would argue
that these synthetic data values are merely draws from the same true
data generating model. In that sense, the observed data is never
replaced, but the population is resampled from the information captured
in the (incomplete) sample. Using this approach, the researcher could
replace the observed data set as a whole with multiple synthetic
versions. Alternatively, the researcher could opt to only replace a
subset of the observed data. For example, one can choose to only replace
dimensions in the data that could be compared with publicly available
data sets or registers. Likewise, synthetisation could be limited to
those values that are disclosive, such as high incomes or high
turnovers.

Conceptually, the synthetic data framework is based upon the building
blocks of multiple imputation of missing data, as proposed by
\citet{rubin_multiple_1987}. Instead of replacing just the missing
values, one could easily replace any observed sensitive values by
multiple draws from the posterior predictive distribution - or a
non-Bayesian predictive distribution. We will refer to both of these as
predictive distributions. Similarly to multiple imputation of missing
data, the multiple synthetic data sets allow for correct statistical
inferences, despite the fact that the analyses do not use the ``true''
value. The analyses over multiple synthetic data sets should be pooled
into a single inference, so that the researcher can draw valid
conclusions from the pooled results. To that respect, the variance
should reflect the added variability that is induced by the imputation
procedure.

Potentially, this approach could fulfil the needs for openly accessible
data, without running into barriers with regard to privacy and
confidentiality constraints. However, there is no such thing as a free
lunch: data collectors have to put effort in creating high-quality
synthetic data. The quality of the synthetic data is highly dependent on
the imputation models, and using flawed imputation models might bias
subsequent analyses
\citep{reiter2004simultaneous, grund2021using, jiang2021balancing}.
Conversely, if the models used to synthesize the data are able to
preserve the relationships between the variables as in the original
data, the synthetic data can be nearly as informative as the observed
data. To fully exploit the benefits of synthetic data, complicating
aspects of creating high-quality synthetic data sets should be kept at a
minimum, while keeping the accuracy of the models as high as possible.

@@

Generally, there are three ways to impute data: joint modeling,
sequential modeling and fully conditional specification
\citep{murray_multiple_2018, ludtke2020regression, grund2021using}. With
joint modeling, the entire joint distribution of the data is specified
as a single multivariate distribution, and the imputations are drawn
from this model. This strategy is implemented in, among others, the
\texttt{R}-package \texttt{jomo} \citep{jomo} to deal with missing data.
However, when the structure of the data increases in complexity,
specifying a single multivariate distribution can become cumbersome. A
more flexible solution has been proposed in sequential regression
imputation, in which the multivariate distribution is separated into
univariate distributions for every variable. Every single variable is
then imputed conditional on a model containing only the variables in the
sequence located before the variable to be imputed. This approach has
been implemented in the R-packages \texttt{mdmb} \citep{mdmb} for
imputation of missing data and \texttt{synthpop} \citep{synthpop} for
imputation of synthetic data.

The \texttt{mice} package \citep{mice} in \texttt{R} \citep{Rproject}
has been developed for multiple imputation to overcome problems related
to nonresponse. In that context, the aim is to replace missing values
with plausible values from the predictive distribution of that variable.
In \texttt{mice}, this is accomplished using fully conditional
specification (FCS) \citep{vanbuuren_fully_2006}, which breaks down the
multivariate distribution of the data
\(\textbf{Y} = (\textbf{Y}_{obs}, \textbf{Y}_{mis})\) into
\(j = 1, 2, \dots, k\) univariate conditional densities, where \(k\)
denotes the number of columns in the data. Using FCS, a model is
constructed for every incomplete variable and the missing values
\(Y_{j, mis}\) are then imputed with draws from the predictive
distribution of \(P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)\) on a
variable-by-variable basis. Note that the predictor matrix \(Y_{-j}\)
may contain yet imputed values from an earlier imputation step, and thus
will be updated after every iteration. This procedure is applied \(m\)
times, resulting in \(m\) completed data sets
\(\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})\),
with \(\textbf{D}^{(l)} = (\textbf{Y}_{obs}, Y^{(l)}_{mis})\).

In \texttt{mice}, the generation of multiply imputed data sets to solve
for unobserved values is straightforward. The following pseudocode
details multiple imputation of any data set that contains missingness
into the object \texttt{imp} with \texttt{m\ =\ 10} imputated sets and
\texttt{maxit\ =\ 7} iterations for the algorithm to converge, using the
default imputations methods.

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(mice)}
\NormalTok{imp }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(data.with.missings, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{7}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

The imputation approach in \texttt{mice} can be easily extended to
generate synthetic values. Rather than imputing missing data, observed
values are then replaced by synthetic draws from the predictive
distribution. For simplicity, assume that the data is completely
observed (i.e., \(\textbf{Y} = \textbf{Y}_{obs}\)). Following the
notation of \citet{reiter_raghunathan_multiple_2007}, let, given \(n\)
observations, \(Z_i = 1\) if any of the values of unit
\(i = 1, 2, \dots, n\), are to be replaced by imputations, and
\(Z_i = 0\) otherwise, with \(\textbf{Z} = (Z_1, Z_2, \dots, Z_n)\).
Accordingly, the data consists of values that are to be replaced and
values that are to be kept (i.e.,
\(\textbf{Y} = (\textbf{Y}_{rep}, \textbf{Y}_{nrep})\). Now, instead of
imputing \(\textbf{Y}_{mis}\) with draws from the predictive
distribution of \(P(Y_{j, mis} | \textbf{Y}_{obs}, \theta)\) as in the
missing data case, \(\textbf{Y}_{rep}\) is imputed from the posterior
distribution of
\(P(Y^{(l)}_{j, rep} | \textbf{Y}^{(l)}_{-j}, \textbf{Z}, \theta)\),
where \(l\) is an indicator for the synthetic data set
(\(l = 1, 2, \dots, m\)). The synthetic values that are imputed at an
earlier step can be used for imputing variable \(j\). This process
results in the synthetic data
\(\textbf{D} = (\textbf{D}^{(1)}, \textbf{D}^{(2)}, \dots, \textbf{D}^{(m)})\).
Note that, given a completely observed data set, iterating over the
predictive distribution is not required.

Hence, creating synthetic values for a completely observed data set into
the object \texttt{syn}, given the same imputation parameters as
specified above, can be realized as follows:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{library}\NormalTok{(mice)}
\NormalTok{imp }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(data.to.synthesize, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{1}\NormalTok{,}
            \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{,}
                           \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(data.to.synthesize),}
                           \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(data.to.synthetize)))}
\end{Highlighting}
\end{Shaded}

Here, the argument \texttt{where} requires a matrix of the same
dimensions as the data, (i.e., a \(n \times k\) matrix) containing
logicals \(z_{ij}\) that indicate which cells are selected to have their
values replaced by draws from the predictive distribution. This approach
allows to replace a subset of the observed data (e.g., by only
specifying those cells that are to be replaced as \texttt{TRUE} in the
\texttt{where}-matrix, leaving the rest to \texttt{FALSE}), or as in the
aforementioned example, the observed data as a whole, resulting in a
data set that partially or completely consists of synthetic data values.
Note that because the data is completely observed, iterating over the
predictive distribution is not required.

If the goal is to create synthetic versions of a data set that suffers
from missingness, the imputation procedure becomes more complex. A
two-step imputation procedure has been proposed by
\citet{reiter2004simultaneous}. In the first step, the missing data is
imputed according to the principles of multiple imputation for missing
data, while in a second step, the synthetic data is imputed conform the
corresponding procedure. Ideally, these steps could be combined using
\texttt{mice}, conditioning the imputations on both the missingness and
the values that are to be synthesized. In practice, replacing observed
and missing cells in a data set by synthetic values into the object
\texttt{syn}, given the same imputation parameters as before, can be
realized by the following code execution.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{syn }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(data.with.missings.to.synthesize, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{7}\NormalTok{, }
            \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{, }
                           \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(data.with.missings.to.synthesize),}
                           \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(data.with.missings.to.synthesize)))}
\end{Highlighting}
\end{Shaded}

However, because the validity of this approach when the data suffers
from missingness has not been investigated yet, we restrict ourselves in
the following sections to a completely observed data set.

Choosing an adequate imputation model to impute the data is paramount,
as a flawed imputation model may drastically impact the validity of
inferences \citep{grund2021using, jiang2021balancing}. Imputation models
should be as flexible as possible to capture most of the patterns in the
data, and to model possibly unanticipated data characteristics
\citep{murray_multiple_2018, rubin_18years_1996}. Generally, multiple
methods can be feasible, including joint modeling, sequential regression
and FCS, all with different parameter specifications, depending on the
imputation problem at hand
\citep{ludtke2020regression, yucel2018sequential, murray2016multiple}.
Various simulations have shown that the most accurate imputation
procedure may vary over different imputation problems
\citep{hanne_review_2021}. Hence, a one-size-fits-all solution does
probably not exist.

However, one of the methods that has shown a reasonable performance over
multiple simulations uses FCS in combination with classification and
regression trees
\citep[CART;][]{breiman_cart_1984, reiter_cart_2005, burgette_reiter_cart_2010, doove_buuren_recursive_2014, raab_practical_2016}.
Due to its non-parametric nature, CART allows to model more complex
patterns (e.g., interactions and non-linear relationships) in the data.
Loosely speaking, CART sequentially splits the predictor space into
non-overlapping regions in such a way that the within-region variance is
as small as possible after every split. As such, CART does not impose
any parametric distribution on the data, making it a widely applicable
method that allows for a large variety of relationships within the data
\citep{islr_2013}. Given these appealing characteristics and the call
for the use of flexible methods when multiply imputing data, we will
focus our illustrations and evaluations of \texttt{mice} to method
\texttt{mice.impute.cart()}, realized by:

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{syn }\OtherTok{\textless{}{-}} \FunctionTok{mice}\NormalTok{(data.to.synthesize, }
            \AttributeTok{m =} \DecValTok{10}\NormalTok{,}
            \AttributeTok{maxit =} \DecValTok{1}\NormalTok{, }
            \AttributeTok{method =} \StringTok{"cart"}\NormalTok{,}
            \AttributeTok{where =} \FunctionTok{matrix}\NormalTok{(}\ConstantTok{TRUE}\NormalTok{, }
                           \AttributeTok{nrow =} \FunctionTok{nrow}\NormalTok{(boys),}
                           \AttributeTok{ncol =} \FunctionTok{ncol}\NormalTok{(boys)))}
\end{Highlighting}
\end{Shaded}

In a nutshell, the above code shows the simplicity of creating
\(m = 10\) synthetic data sets using \texttt{mice}. In practice,
however, one should take some additional complicating factors into
account. For example, one should account for deterministic relations in
the data. Additionally, relations between variables may be described
best using a different model than \texttt{CART}. Such factors are data
dependent and should be considered by the imputer. In the next section,
we will describe how the \texttt{mice::boys} data
{[}fredriks\_boys\_2000{]} can be adequately synthesized. Additionally,
we will show through simulations that this approach yields valid
inferences.

@@

Usually, one would draw samples from a population that can be
synthesized to evaluate the performance of the synthetisation methods.
As we only have access to a single sample, 1000 bootstrap samples of the
\texttt{boys} data have been synthesized with \(m = 5\) imputations for
every data cell to induce an appropriate amount of sampling variance.
This approach precludes us from knowing the \emph{true} data-generating
model, and in this sense provides a more stringent test of the
applicability of \texttt{mice} for real-life synthetisation procedures.
Specifically, every bootstrapped sample is treated as an actual sample
from a population that is synthesized using \texttt{mice}, using a model
that is built to approximate the true data-generating mechanisms as
closely as possible. As relationships in realistic data are generally
more complex than relationships in parametrically simulated data,
achieving good performance on a real data set is likely to be more
indicative of practical applicability than good performance on a
parametrically simulated data set.

In the simulations, synthetic values are generated using the
\texttt{CART} imputation method for all columns, except for
\texttt{bmi}. The deterministic relation \texttt{bmi} which will be
synthesized passively based on the synthetic values for \texttt{hgt} and
\texttt{wgt} to preserve the relation in the synthetic data. Additional
parameters that come with the use of \texttt{mice.impute.cart()} are the
complexity parameter \texttt{cp} and the minimum number of observations
in any terminal node \texttt{minbucket}, that both constrain the
flexibility of the imputation model. The values of the parameters
\texttt{cp} and \texttt{minbucket} ought to adhere to the call for
imputation models that are as flexible as possible. Appropriate values
for these parameters, as well as the input for the
\texttt{predictorMatrix}, depend on the data at hand. In the current
example, the complexity parameter is specified at \texttt{cp\ =\ 1e-08}
rather than the default value \texttt{1e-04}, and the minimum number of
observations in each terminal node is set at \texttt{minbucket} \(= 3\)
rather than the default value \(5\). By allowing for more complexity in
the imputation model, bias in the estimates from the synthetic data set
is reduced. Additionally, since the synthesis pattern is monotone, the
number of iterations can be set to \texttt{maxit\ =\ 1}
\citep[e.g.,][Ch. 3]{drechsler_synthetic_2011}.

@@

Obtaining a final point estimate of the parameter of interest \(Q\)
after imputation is fairly easy and no different from pooling in the
case of missing data \citep{rubin_multiple_1987}. One can calculate the
average of the \(m\) point estimates \(q^{(l)}\)

\[
\bar{q}_m = \sum_{l = 1}^m \frac{q^{(l)}}{m},
\] with \(l = 1, \dots, m\).

Also similarly to the missing data case, variances, and subsequently
confidence intervals, should incorporate the increase in variance that
is due to imputation
\citep{reiter_partially_inference_2003, drechsler_synthetic_2011}. Yet,
the increase in variance due to imputation differs according to whether
missing values are imputed or observed data is replaced by synthetic
values. Whereas the variance estimate after imputation of missing data
needs to account for the fact that a certain amount of information in
the data is missing, variance estimation from synthetic data does not
suffer from this issue. The adjusted variance estimate that follows from
using multiple synthetic data sets only suffers from the fact that a
finite number of \(m\) synthetic data sets are used to resemble the
observed data. Hence, the according variance estimate for synthetic data
as developed by \citet{reiter_partially_inference_2003} yields

\[
T = \bar{u}_m + \frac{b_m}{m},
\]

@@

\[
\bar{u}_m = \sum_{l = 1} \frac{u^{(l)}}{m},
\] where \(u^{(l)}\) denotes the variance estimate in the \(l\)th
synthetic data set. These pooling rules are implemented in \texttt{mice}
as the function \texttt{pool.syn}. \citet{reiter_kinney_2012} proved
that synthesizing data from the posterior predictive distribution is not
required for these pooling rules to hold. In fact, the parameters of an
imputation model can be fixed to their maximum likelihood estimates or
their posterior modes, and synthetic values can be drawn after plugging
these values into the imputation models. This observation as been taken
a step further by \citet{raab_practical_2016}, who developed pooling
rules that do not require multiple synthetic data sets and performed
highly similar to the aforementioned pooling rules. However, we adhere
to the originally proposed pooling rules by
\citet{reiter_partially_inference_2003}, who defined the total variance
in terms of sampling variance and between-imputation variance.

@@

The sampling distribution of the means of the continuous variables for
both the bootstrapped and the synthetic data can be found in Figure 1,
while the number of observations per category in the bootstrapped data
and the averaged number of observations per category over the synthetic
data sets can be found in Figure 2.

\includegraphics{Manuscript_files/figure-latex/univariate-statistics-1.pdf}

Both figures show that the marginal distributions of the bootstrapped
data and the synthetic data are highly similar. Additionally, the
univariate descriptives for the original data and the synthetic data can
be found in Table 2.

\includegraphics{Manuscript_files/figure-latex/univariate-statistics-II-1.pdf}

@@

We demonstrate that generating synthetic data sets with \texttt{mice}
using CART in \texttt{R} is a straightforward process that fits well in
a data analysis pipeline. The approach is hardly different from using
multiple imputation to solve problems related to missing data, and hence
can be expected to be familiar to applied researchers. The multiple
synthetic sets yield valid inferences on the true underlying data
generating mechanism, thereby capturing the nature of the original data.
This makes the multiple synthetisation procedure with \texttt{mice}
suitable for further dissemination of synthetic data sets. It is
important to note that in our simulations we used a single iteration and
relied on CART as the imputation method. A single iteration is
sufficient only when the true data is completely observed, or when the
missingness pattern is monotone \citep{drechsler_synthetic_2011}. If
both observed and unobserved values are to be synthesized, then more
iterations and a careful investigation into the convergence of the
algorithm are in order. Synthetic data generation is therein no
different than multiple imputation.

However, when creating synthetic data with \texttt{mice}, close
attention should be paid to three distinct aspects of the data:
incorporating the additional uncertainty that is due synthesizing (part
of) the data, potential disclosure risk that remain after synthesizing
the data, and the utility of the synthetic data. First, the procedure of
generating multiple synthetic sets may seem overly complicated. We would
like to emphasize that analyzing a single synthesized set, while perhaps
unbiased, would underestimate the variance properties that are so
important in drawing statistical inferences from finite data sets. After
all, we are often not interested in the sample at hand, but aim to make
inferences about the underlying data generating mechanism as reflected
in the population. Properly capturing the uncertainty of synthetic data
sets, just like with incomplete data sets, is therefore paramount. To
achieve this, we adopted a bootstrap scheme in our simulations, to
represent sampling from a superpopulation. However, when a single sample
should be the reference, in the sense that the sample reflects the
population one wishes to make inferences about, adjusted pooling rules
are required that are similar to the procedure outlined in
\citet{vink2014pooling}. The corresponding pooling rules have not been
derived yet and the incorporation in data analysis workflows would
require proper attention from developers alike.
